### 2014-11-02 paper/experiments.tex
fixed spelling and grammer mistakes

We conducted several performance tests to measure the throughput and scalability of our system. All these tests were performed in an LAN system called Lattice which has a network bandwidth of 1Gbps. All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 Gbp of memory. We used Apache S4 0.6.0 verson and Apache storm 0.9.2 version. Sample code used for all tests can be found here\cite{solutionCode}.

\subsection{Inter node communication}
We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. Our solution and Yahoo S4\cite{neumeyer2010s4} EventProducer reads a file containing over 7500000 ecg records and push events with multiple threads. Twitter Storm\cite{twitterStorm} does not allow user threads to push data. So we used the Spout thread to send messages into the system. EventReceiver  receives the ECG events, process them and calculate heart rate interval periodically.  
\subsubsection{Throughput}
We execute our system with 1, 2 and 4  EventReceivers for each system and measured the throughput, load average and network bandwidth  for each case. Throughput was measured at the EventProducer calculating total time required to send messages and the total number of messages send. Load average and network bandwidth were measured using top and atop linux commands respectively. Figure \ref{throuput} shows the throughput variation with the number of nodes for all systems.

@@

As shown in the Figure \ref{throuput} our solution out performs Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}. However by looking the Figure \ref{throuput} one might think even our solution does not scalable although it perform high. In order to examine the reason behind this we looked into the network bandwidth and CPU load average values.

@@

By looking at the graphs, first it can be observed that our solution utilities all available network bandwidth (98\%) even with two receivers. In only one receiver case, high CPU load average at receiver indicates it has utilized available CPU power. Adding two receivers has increased the CPU power at receiving side utilizing all available network bandwidth. Therefore it can not increase the throughput without increasing network bandwidth although more CPU power available at receiver nodes. Having receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver since load is shared among the receivers. However for other two systems, it neither hits the maximum network bandwidth available nor the full CPU power available at any stage. If a system does not make use of the available resources then adding more resources won't scale up the system. As explained in earlier, our solution utilities all available resources by using parallel tcp connections and having an efficient 
message parsing technique.

\subsubsection{Efficiency of Message Serialization}
After measuring the throughput we examined the  total amount of extra bytes each system sends to transfer the information from EventProducer to EventReceiver. Our ECGEvent has two double fields called time and value (ecg signal value) and a streamID which is a 4 byte string. So we calculated total bytes to send this information as 20 bytes (16 for two doubles and 4 for streamID). Then by using the throughput and the bandwidth usage of the system we calculated the total number of bytes each system uses at network layer to send this message. By reducing 20 information bytes we can calculate the over head bytes for each system. Figure \ref{efficiency} compares these numbers.

@@

As shown in the figure \ref{efficiency} Twitter storm\cite{twitterStorm} uses a minimum over head while Yahoo S4\cite{neumeyer2010s4} performs badly compared to other two. This overhead is due to its usage of  java serializations to serialize data. Then we analyzed for what our solution spends extra 32 bytes. Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process and to parse the message at the server. Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and other 4 bytes due to tcp over head. This is the trade off we had to pay to improve the parallelism compared to direct one tcp communication per process.

\subsubsection{Latency}
We examined the message latency between the event producer and receiver. For this experiment we send the time stamp at the producer along the message and calculated the latency at the receiver. We took 3000 samples over experiment duration. As shown in the figure \ref{latancydis}, we could calculate the mean latency as 221.86 with the 95\% confidence interval of (218.18, 225.55).

@@

This application predicts load averages using the previous values and a given machine learning algorithm. We implement this logic using nodes as shown in the Figure \ref{multigraph}. First producer reads the data file and send events to the avgCal node which calculate the last minite average and send the same event to both plugPredict and housePredict processors. Both plugPredict and housePredict processors predict the next values and send events to receivers. The original problem only requires to send those prediction events in 30s intervals. But we send a prediction event for each messages to observe how the system works with high load.

@@

For one receiver unit case, we observed very high load average and network bandwidth usage at avgCal node since it send messages to two nodes. Then as shown in the figure \ref{scalability} we could linerly scale up the system by adding more receiver units to add more cpu power to system. When increasing the receiver units, we can observe throughput of the system increases proportional to number of receiving units. CPU load average of 13 at the producer indicates it has utilized all available CPU power and we need to add more producers to scale this system further up.

>>>>

We conducted several performance tests to measure the throughput and scalability of our system. All these tests were performed in an LAN system called Lattice which has a network bandwidth of 1Gbps. All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 Gbp of memory. We used Apache S4 0.6.0 version and Apache storm 0.9.2 version. Sample code used for all tests can be found here\cite{solutionCode}.

\subsection{Inter node communication}
We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. Our solution and Yahoo S4\cite{neumeyer2010s4} EventProducer reads a file containing over 7500000 ecg records and push events with multiple threads. Twitter Storm\cite{twitterStorm} does not allow user threads to push data. So we used the Spout thread to send messages into the system. EventReceiver  receives the ECG events, process them and calculate heart rate interval periodically.  
\subsubsection{Throughput}
We execute our system with 1, 2 and 4  EventReceivers for each system and measured the throughput, load average and network bandwidth  for each case. Throughput was measured at the EventProducer by calculating the total time required to send messages and the total number of messages send. Load average and network bandwidth were measured using top and atop linux commands respectively. Figure \ref{throuput} shows the throughput variation with the number of nodes for all systems.

@@

As shown in the Figure \ref{throuput} our solution outperforms Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}. However by looking the Figure \ref{throuput} one might think even our solution does not scalable although it perform high. In order to examine the reason behind this we looked into the network bandwidth and CPU load average values.

@@

By looking at the graphs, first it can be observed that our solution utilities all available network bandwidth (98\%) even with two receivers. In only one receiver case, high CPU load average at receiver indicates it has utilized available CPU power. Adding two receivers has increased the CPU power at receiving side utilizing all available network bandwidth. Therefore it can not increase the throughput without increasing network bandwidth although more CPU power available at receiver nodes. Having receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver since load is shared among the receivers. However for other two systems, it neither hits the maximum network bandwidth available nor the full CPU power available at any stage. If a system does not make use of the available resources then adding more resources won't scale up the system. As explained in earlier, our solution utilities all available resources by using parallel tcp connections and having an efficient 
message parse technique.

\subsubsection{Efficiency of Message Serialization}
After measuring the throughput we examined the  total amount of extra bytes each system sends to transfer the information from EventProducer to EventReceiver. Our ECGEvent has two double fields called time and value (ecg signal value) and a streamID which is a 4 byte string. So we calculated total bytes to send this information as 20 bytes (16 for two doubles and 4 for streamID). Then by using the throughput and the bandwidth usage of the system we calculated the total number of bytes each system uses at network layer to send this message. By reducing 20 information bytes we can calculated the overhead bytes for each system. Figure \ref{efficiency} compares these numbers.

@@

As shown in the figure \ref{efficiency} Twitter storm\cite{twitterStorm} uses a minimum over head while Yahoo S4\cite{neumeyer2010s4} performs badly compared to other two. This overhead is due to its usage of  java serialization to serialize data. Then we analyzed for what our solution spends extra 32 bytes. Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process and to parse the message at the server. Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and another 4 bytes due to tcp overhead. This is the trade off we had to pay to improve the parallelism compared to direct one tcp communication per process.

\subsubsection{Latency}
We examined the message latency between the event producer and receiver. For this experiment we sent the timestamp at the producer along the message and calculated the latency at the receiver. We took 6000 samples over experiment duration. As shown in the figure \ref{latancydis}, we could observe the mean latency as 185.05 with the 95\% confidence interval of (182.56 187.54).

@@

This application predicts load averages using the previous values and a given machine learning algorithm. We implement this logic using nodes as shown in the Figure \ref{multigraph}. First producer reads the data file and send events to the avgCal node which calculate the last minute average and send the same event to both plugPredict and housePredict processors. Both plugPredict and housePredict processors predict the next values and send events to receivers. The original problem only requires to send those prediction events in 30s intervals. But we sent a prediction event for each messages to observe how the system works with high load.

@@

For one receiver unit case, we observed very high load average and network bandwidth usage at avgCal node since it sends messages to two nodes. Then as shown in the figure \ref{scalability} we could linearly scale up the system by adding more receiver units to add more cpu power to system. When increasing the receiver units, we can observe throughput of the system increases proportional to number of receiving units. CPU load average of 13 at the producer indicates it has utilized all available CPU power and we need to add more producers to scale this system further up.

### 2014-11-02 paper/relatedwork.tex
fixed spelling and grammer mistakes

Map reduce\cite{dean2008mapreduce} is a widely adapted technology to process batch data with the populariy of Apache Hadoop\cite{hadoop}. In a map reduce envirionment a problem is partition into smaller parts identified by a key and process parallely. Although these systems support user defined functions, data processing flow of these systems are fixed. Ciel\cite{murray2011ciel} address this problem by dynamically generating dataflow graph. Since all these systems communicate through file system they inherently not suitable for real time processing\cite{lam2012muppet}.

Stream Processing Core (SPC)\cite{Amini2006}, Yahoo S4\cite{neumeyer2010s4} and Twitter Storm\cite{twitterStorm} address the above issue by supporting direct communication. These systems are based on the Actor model\cite{agha1985actors} and each system has a notion of processing element or computation. Processing Elements are used to perform user defined logic on the receiving events and emit newly generated events to other processing elements. In addition to basic communication these systems provide fault tolarance features such as relaiable message delivary and dynamic membership management. However most of these system does not focus on improving internode communication performance.

Spark\cite{zaharia2010spark} introduces its RRDs\cite{zaharia2012resilient} to achive better fault tolarance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically. Spark Streaming\cite{zaharia2013discretized} extends this concept to distributed stream processing with the concept of discretized events. Spark streaming\cite{zaharia2013discretized} processes data as batches while storing them as RDDs \cite{zaharia2012resilient} to achive higher throughput. This batch processing may introduce latances which are not acceptable for some applications. Further it is not clear how to implement complex event processing algorithumns such as we use our bench marks with Spark Streaming. MillWheel\cite{akidau2013millwheel} is the stream processing system used at Google to process web queries. MillWheel\cite{akidau2013millwheel} provides relaiable message processing as well as state persistance for its graph based computations. Unlike in our system MillWheel treat messages as binary 
messages leaving message parsing and serialization to application layer.

>>>>

Map reduce\cite{dean2008mapreduce} is a widely adapted technology to process batch data with the popularity of Apache Hadoop\cite{hadoop}. In a map reduce environment a problem is partition into smaller parts identified by a key and process parallely. Although these systems support user defined functions, data processing flow of these systems are fixed. Ciel\cite{murray2011ciel} address this problem by dynamically generating data flow graph. Since all these systems communicate through file system they inherently not suitable for real time processing\cite{lam2012muppet}.

Stream Processing Core (SPC)\cite{Amini2006}, Yahoo S4\cite{neumeyer2010s4} and Twitter Storm\cite{twitterStorm} address the above issue by supporting direct communication. These systems are based on the Actor model\cite{agha1985actors} and each system has a notion of processing element or computation. Processing Elements are used to perform user defined logic on the receiving events and emit newly generated events to other processing elements. In addition to basic communication these systems provide fault tolarance features such as reliable message delivery and dynamic membership management. However most of these system does not focus on improving internode communication performance.

Spark\cite{zaharia2010spark} introduces its RRDs\cite{zaharia2012resilient} to achieve better fault tolerance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically. Spark Streaming\cite{zaharia2013discretized} extends this concept to distributed stream processing with the concept of discretized events. Spark streaming\cite{zaharia2013discretized} processes data as batches while storing them as RDDs \cite{zaharia2012resilient} to achieve higher throughput. This batch processing may introduce latencies which are not acceptable for some applications. Further it is not clear how to implement complex event processing algorithms such as we use our benchmarks with Spark Streaming. MillWheel\cite{akidau2013millwheel} is the stream processing system used at Google to process web queries. MillWheel\cite{akidau2013millwheel} provides reliable message processing as well as state persistance for its graph based computations. Unlike in our system MillWheel treat messages as binary 
messages leaving message parsing and serialization to application layer.

### 2014-11-02 paper/system_architecture.tex
fixed spelling and grammer mistakes

\section{System Design}
In this section we elaborate the underling system design which is capable of processing 2.4 million messages per second. We emphazise three features of this design. Firstly between any two nodes there is a set of tcp connections (typically 20 connections)  to send messages. All processes share the same set of connections which we called as the bridge. The advantage of this approach compared to having tcp connections per process is that it utilizes all available bandwidth even there is only one process per node. However since this is a common bridge now we need to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages. As we show latter our experiments shows this approach perform several times better than the other systems dispite above overhead. Secondly we have implemented OutputStream and InputStream interfaces on top of java.nio API. Using these input and output streams we provide DataOutput and DataInput APIs to directly serialize message data to binary format without sending any metadata along the message. Finally except messages itself all the other objects used to serialize the messages and parse the messages are not created at message processing time. When a process sends the first message all these objects are created and it is a one time cost. Rest of this section further explain design techniques we used to realise these concepts.
\subsection{Inter Process Communication}
Figure \ref{interprocess} shows the design involves in sending a message from one process to another. Lets assume there is a message at a client side process and that process wants to send this message to another process. First client process sends this message to its ElementContainer. ElementContainer holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams. Stream decides which node to send this message and passes the message along the taget node details to ConnectionManager which holds ClientConnections for each node. ConnectionManager picks the correct ClientConnection for target node and sends the message to ClientConnection. ClientConnection picks an available DataOutput from its pool of DataOutputs and invokes the serialize method of the message to send that to server side.

At the server side there is a set of ServerTasks which read receiving messages from pool of DataInput streams available in the ServerConnection (we register these connections at the connections creating time). Once a server task receives a binary message, it creates the event using the sending process id to identify the event type. Then it pass this message to the WorkerContainer which holds all processes. Finally WorkerContainer dispatch this message to correct process uisng receiving process id. Message ordering happens at the process level if required. This elastrates how we have shared a pool of tcp connections among process and next we show how we have implemented java.io.OutputStream and java.io.InputStream with java.nio API without  creating objects dynamically.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}
\subsection{Client Side}
Figure \ref{client} shows how we have implemented java.io.OutputStream interface using our DataWriter. DataWriter contains a ByteBuffer. When a higher layer method writes data to DataOuput, DataWriter put those bytes to its byte buffer. When a selection event occurs, Datawriter writes its byte buffer to underline SocketChannel. This way we use one byte buffer as an intermediatory to create an output stream with java.nio API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{client.png}
        \caption{OutputStream Implementation}
        \label{client}
\end{figure}
\subsection{Server Side}
Figure \ref{server} shows how we have implemented java.io.InputStream interface using our DataReader. Similar to clinet side, DataReader has a java.nio.ByteBuffer. When a selection event occurs DataReader input bytes from the SocketChannel to its byte buffer. When a ServerTask reads data through DataInput, it reads data from the byte buffer and passes to higher layer. In this way we have implemented an input stream with java.nio API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{server.png}
        \caption{InputStream Implementation}
        \label{server}
\end{figure}

>>>>

\section{System Design}
In this section we elaborate the underlying system design which is capable of processing 2.4 million messages per second. We emphasize three features of this design. Firstly between any two nodes there is a set of tcp connections (typically 20 connections)  to send messages. All processes share the same set of connections which we called as the bridge. The advantage of this approach compared to having tcp connections per process is that it utilizes all available bandwidth even there is only one process per node. However since this is a common bridge now we need to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages. As we show later our experiments shows this approach perform several times better than the other systems despite above overhead. Secondly we have implemented OutputStream and InputStream interfaces on top of java.nio API. Using these input and output streams we provide DataOutput and DataInput APIs to directly serialize message data to binary format without sending any metadata along the message. Finally except messages itself all the other objects used to serialize the messages and parse the messages are not created at message processing time. When a process sends the first message all these objects are created and it is a one time cost. Rest of this section further explain design techniques we used to realise these concepts.
\subsection{Inter Process Communication}
Figure \ref{interprocess} shows the design involves in sending a message from one process to another. Lets assume there is a message at a client side process and that process wants to send this message to another process. First client process sends this message to its ElementContainer. ElementContainer holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams. Stream decides which node to send this message and passes the message along the target node details to ConnectionManager which holds ClientConnections for each node. ConnectionManager picks the correct ClientConnection for target node and sends the message to ClientConnection. ClientConnection picks an available DataOutput from its pool of DataOutputs and invokes the serialize method of the message to send that to server side.

At the server side there is a set of ServerTasks which read receiving messages from pool of DataInput streams available in the ServerConnection (we register these connections at the connections creating time). Once a server task receives a binary message, it creates the event using the sending process id to identify the event type. Then it pass this message to the WorkerContainer which holds all processes. Finally WorkerContainer dispatch this message to correct process using receiving process id. Message ordering happens at the process level if required. This illustrates how we have shared a pool of tcp connections among process and next we show how we have implemented OutputStream and InputStream with non blocking I/O API without  creating objects dynamically.
\begin{figure}[!tii]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}
\subsection{Client Side}
Figure \ref{client} shows how we have implemented OutputStream interface using our DataWriter. DataWriter contains a ByteBuffer. When a higher layer method writes data to DataOuput, DataWriter put those bytes to its byte buffer. When a selection event occurs, Datawriter writes its byte buffer to underline SocketChannel. This way we use one byte buffer as an intermediary to create an output stream with non blocking I/O API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{client.png}
        \caption{OutputStream Implementation}
        \label{client}
\end{figure}
\subsection{Server Side}
Figure \ref{server} shows how we have implemented InputStream interface using our DataReader. Similar to client side, DataReader has a ByteBuffer. When a selection event occurs DataReader input bytes from the SocketChannel to its byte buffer. When a ServerTask reads data through DataInput, it reads data from the byte buffer and passes to higher layer. In this way we have implemented an input stream with non blocking I/O API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{server.png}
        \caption{InputStream Implementation}
        \label{server}
\end{figure}

### 2014-11-02 paper/system_overview.tex
fixed spelling and grammer mistakes

\section{System Overview}
Our system mainly consists of a set of processes which receive events, process them and forward to next processes. We use the term adapters for special processes which receive events from out side systems. These processes and adapters forms a graph which represents the way a given event stream being processed within the system. Our current implementation does not support fault tolarance and have adapted topoloy concept from twitter storm\cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}. Rest of this section further describes this concept at a user level. In the next section we provide a detail explanation about how we have make this interprocess communication efficient. 
\subsection{Process Graph}
A Process graph (Figure \ref{processgraph}) mainly consists of the adapters, processes which does the processing logic and their interaction patterns. Adapters receive events from out side systems and forward them to other processes. Processes process events according to a given logic and emits the generated events either to other processes or to out side systems.

@@

 The Event interface has a method to get the key for a given event. This key is used in sending messages to next processes as described earlier. Parse and serialize methods of event interface are used to direcly serialize the event attributes to underline streams. When implementing these methods deveopers can define attibute serializing order in serialize method and use same order at the parse method to avoid meta data passing with the binary format. Process interface has a method called onEvent which underline framework invokes when it receives an event to that process. The start method of the adapter interface is invoked by the framework and that can be used to pull events from out side systems onces the system starts. At the initilization time both processes and adapters receive the container which can be used to send events to other processes.

@@

\subsection{Application Developement}

An application specifies the runtime graph to process a pirticular event stream. Therefore an appication mainly consists of a set of processes, adapters and events and their interactions. Processes, adapters and events can be developed by implementing the respective interfaces as shown above. Then the interactions can be specified using a json file.

>>>>

\section{System Overview}
Our system mainly consists of a set of processes which receive events, process them and forward to next processes. We use the term adapters for special processes which receive events from outside systems. These processes and adapters form a graph which represents the way a given event stream is processed within the system. Our current implementation does not support fault tolerance and has adapted topology concept from twitter storm\cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}. Rest of this section further describes this concept at a user level. In the next section we provide a detailed explanation about how we have made this inter-process communication efficient. 
\subsection{Process Graph}
A Process graph (Figure \ref{processgraph}) mainly consists of the adapters, processes which does the processing logic and their interaction patterns. Adapters receive events from out side systems and forward them to other processes. Processes process events according to a given logic and emits the generated events either to other processes or to outside systems.

@@

 The Event interface has a method to get the key for a given event. This key is used in sending messages to next processes as described earlier. Parse and serialize methods of event interface are used to directly serialize the event attributes to underline streams. When implementing these methods developers can define attribute serialization order in serialize method and use same order at the parse method to avoid meta data passing with the binary format. Process interface has a method called onEvent which underline framework invokes when it receives an event to that process. The start method of the adapter interface is invoked by the framework and that can be used to pull events from outside systems once the system starts. At the initialization time both processes and adapters receive the container which can be used to send events to other processes.

@@

\subsection{Application Development}

An application specifies the runtime graph to process a particular event stream. Therefore an application mainly consists of a set of processes, adapters and events and their interactions. Processes, adapters and events can be developed by implementing the respective interfaces as shown above. Then the interactions can be specified using a json file.
